---
title: "Taylor HT Data Analysis"
author: "Taylor Ely and Zack Gold"
date: "12/04/2020"
output: html_document
---

#Load Packages
```{r}
library(ranacapa)
library(phyloseq)
library(tidyverse)
library(dplyr)
library(tibble)
library(reshape2)
library(ggplot2)
library(treemapify)
library(vegan)
library(plotly)
library(optparse)
library("qiimer")
library(ggrepel)
library(wesanderson) 
library(cluster)
library(biomformat)
library('gridBase')
library(devtools)
library(grid)
library(gridExtra)
library(ape)
library(scales)
library(tidyr)
library(RColorBrewer)
library(treemap)
library(iNEXT)
library(stats)
```
#Load All Data Sets
```{r}

#All Species eDNA Index
t_index <- read.csv("ASV_taylor_edna_index_no_mammals.csv")

#Species Monitored by NPS, PISCO, and Reef Check
monitored <- read.csv("ASV_taylor_edna_index_monitored_only.csv")

#Species Monitored by Only Reef Check
reefcheck <- read.csv("ASV_taylor_edna_index_reefcheck_only.csv")

#Metadata
metadata <- read.table("MetaData_eDNA_degradation2020.txt", header = 1, sep = "\t", stringsAsFactors = F)


```
#Figure 1. Detection of Foreign eDNA over Time
```{r}

#Select Grass Carp Species Only
t_index %>% as.data.frame() %>% 
  filter(., sum.taxonomy =="Eukaryota;Chordata;Actinopteri;Cypriniformes;Cyprinidae;Ctenopharyngodon;Ctenopharyngodon idella")-> gcarp_idx

# Convert to Long Data
columns <- colnames(t_index)
remove <- c("sum.taxonomy")
gathercols <-  columns[! columns %in% remove] 

gcarp_idx_clean <- gather(gcarp_idx, e_index, gathercols, factor_key=TRUE)

#Remove Name
gcarp_idx_clean[1:36,] -> gcarp_idx_clean

#Format Data
gcarp_idx_clean %>% 
  rename(., original_name =e_index, e_index=gathercols) -> gcarp_idx_clean

gcarp_idx_clean$e_index <- as.numeric(gcarp_idx_clean$e_index)

gcarp_idx_clean %>% 
  left_join(metadata) %>% 
  filter(., Names != "T0") ->eDNA_data


#Generate Colors
wes_palette("Darjeeling1") -> col_darjeeling1
col_species<-c(col_darjeeling1[2],col_Cavalcanti1[2], col_Darjeeling2[2])
col_nmds<-c(col_Zissou1[1],col_Cavalcanti1[2],col_Darjeeling2[3])
wes_palette("IsleofDogs1") -> col_dogs1
wes_palette("Cavalcanti1") -> col_Cavalcanti1
wes_palette("BottleRocket2") -> col_BottleRocket2
wes_palette("BottleRocket1") -> col_BottleRocket1
wes_palette("Royal1") -> col_Royal1
wes_palette("Royal2") -> col_Royal2
wes_palette("Zissou1") -> col_Zissou1
wes_palette("Darjeeling2") -> col_Darjeeling2
wes_palette("Rushmore1") -> col_Rushmore1

#Plot Grass Carp eDNA Index vs. Time after Foreign eDNA Introduction 
jpeg("Figure1.tiff", width=5000, height=3300, res=600)
timepoint<- c(1.5,1.5,1.5,3,3,3,4.5,4.5,4.5,6,6,6,7.5,7.5,7.5,9,9,9,10.5,10.5,10.5,12,12,12,24,24,24)
grasscarp_conc<-eDNA_data$e_index
grasscarp_conc2<-grasscarp_conc[c(-1,-2,-3,-7,-8,-9,-10,-11,-12)] # remove before and after times
plot(timepoint,grasscarp_conc2,pch="", xlab = "Time after Foreign eDNA Introduction (hr)", ylab=" ", cex.lab=1.3, cex.axis=1.3,  ylim=c(0,0.4), axes=FALSE)+ axis(1, at = seq(0, 25, by = 1.5)) + axis(2, at = seq(0, 0.4, by = 0.1))+ title(ylab="eDNA Index", line=2.5, cex.lab=1.3)
points(timepoint,grasscarp_conc2, pch=0:2, cex=2.5)
legend(20, 0.4, legend=c("Site A", "Site B", "Site C"),
       pch = 0:2, cex=1.4)
dev.off()

```

#Ranacapa Code
```{r anacapa to phyloseq functions from ranacapa library imported manually since they do not work on R3.6.0,  results="hide", warning=FALSE, include=FALSE}
#' Takes a site-abundance table from Anacapa, and summarizes to each unique taxon in the sum.taxonomy column
#' @param taxon_table OTU table from Anacapa
#' @author Gaurav Kandlikar
#' @examples
#' good_taxon_table <- data.frame(sum.taxonomy = c("a;b;c;d;f;u", "p;q;r;s;t;u"),
#' site_1 = c(0,1), site_2 = c(10, 20))
#' group_anacapa_by_taxonomy(good_taxon_table)
#' @export
group_anacapa_by_taxonomy <- function(taxon_table) {
  # If tables were made in New Anacapa (with Domain), get rid of domain...
  if(stringr::str_count(taxon_table$sum.taxonomy, ";")[1] == 6) {
    taxon_table$sum.taxonomy = gsub(pattern = "^.*?;", replacement = "", x = taxon_table$sum.taxonomy)
  }
  taxon_table %>%
    dplyr::filter(sum.taxonomy != "") %>%
    dplyr::group_by(sum.taxonomy) %>%
    dplyr::summarize_if(is.numeric, sum) %>%
    dplyr::mutate(sum.taxonomy = as.character(sum.taxonomy)) %>%
    data.frame
}


#' Takes an site-abundance table from Anacapa, along with a qiime-style mapping file, and returns a phyloseq object
#' @param taxon_table Taxon table in the anacapa format
#' @param metadata_file Metadata file with rows as sites, columns as variables
#' @return phyloseq class object
#' @author Gaurav Kandlikar
#' @examples
#' good_taxon_table <- data.frame(sum.taxonomy = c("a;b;c;d;f;u", "p;q;r;s;t;u"),
#' site_1 = c(0,1), site_2 = c(10, 20))
#' good_maps <- data.frame(site = c("site_1", "site_2"), season = c("wet", "dry"),
#'  host = c("oak", "sage"))
#' convert_anacapa_to_phyloseq(good_taxon_table, good_maps)
#' @export

convert_anacapa_to_phyloseq <- function(taxon_table, metadata_file) {

  # Validate the files
  validate_input_files(taxon_table, metadata_file)

  # Group the anacapa ouptut by taxonomy, if it has not yet happened, and turn it into a matrix
  taxon_table2 <- group_anacapa_by_taxonomy(taxon_table) %>%
    tibble::column_to_rownames("sum.taxonomy") %>%
    as.matrix
  # Reorder the columns (sites) for ease of displaying later
  taxon_table2 <- taxon_table2[ , order(colnames(taxon_table2))]

  # Convert the matrix into a phyloseq otu_table object, with taxa as the rows
  ana_taxon_table_physeq <- phyloseq::otu_table(taxon_table2, taxa_are_rows = TRUE)

  # Extract the rownames of the matrix above- this has the full taxonomic path.
  # Split the taxonomic path on semicolons, and turn the resulting matrix into
  # a phyloseq tax_table object
  taxon_names <- reshape2::colsplit(rownames(taxon_table2), ";",
                          names = c("Phylum","Class","Order","Family","Genus","Species")) %>%
    as.matrix
  rownames(taxon_names) <- rownames(taxon_table2)

  tax_physeq <- phyloseq::tax_table(taxon_names)
  colnames(tax_physeq) <- c("Phylum","Class","Order","Family","Genus","Species")

  # Make a phyloseq object out of the otu_table and the tax_table objects
  physeq_object <- phyloseq::phyloseq(ana_taxon_table_physeq, tax_physeq)

  # Make sure the mapping file (ie the site metadata) is ordered according to site name
  rownames(metadata_file) <- metadata_file[, 1]
  metadata_file <- metadata_file[order(metadata_file[, 1]), ]

  # Convert the mapping file into a phyloseq sample_data object, and merge it with the
  # phyloseq object created above to make a phyloseq object with otu table, tax table, and sample data.
  sampledata <- phyloseq::sample_data(metadata_file)
  phyloseq::merge_phyloseq(physeq_object, sampledata)
}

#' Takes a phyloseq object with an otu_table object and returns a vegan style community matrix.
#' @param physeq_object phyloseq object with an otu_table object within
#' @return vegan-style community matrix
#' @examples
#' good_taxon_table <- data.frame(sum.taxonomy = c("a;b;c;d;f;u", "p;q;r;s;t;u"),
#' site_1 = c(0,1), site_2 = c(10, 20))
#' good_maps <- data.frame(site = c("site_1", "site_2"), season = c("wet", "dry"),
#' host = c("oak", "sage"))
#' physeq_object <- convert_anacapa_to_phyloseq(good_taxon_table, good_maps)
#' vegan_otu(physeq_object)
#' @export
vegan_otu <- function(physeq_object) {
  OTU <- phyloseq::otu_table(physeq_object)
  if (phyloseq::taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(methods::as(OTU, "matrix"))
}

#' Remove "xxx_seq_number" column from ana_taxon_table file if it exists
#' takes one taxon table as its input, and if it include
#' a column named "xxx_seq_number", it gets rid of that column - it's not of use to us
#' any longer
#'
#' @param taxon_table taxonomy table from Anacapa
#' @return ana_taxon_table file, with "xxx_seq_number" column removed (if it existed)
#' @examples
#' good_taxon_table <- data.frame(seq_number = c(1,2),
#' sum.taxonomy = c("a;b;c;d;f;u", "p;q;r;s;t;u"),
#' site_1 = c(0,1), site_2 = c(10, 20))
#' scrub_seqNum_column(good_taxon_table)
#' @export
scrub_seqNum_column <- function(taxon_table) {
  to_return <- taxon_table %>% dplyr::select(-dplyr::matches("seq_number"))
  return(to_return)
}

#' Replace empty calls in Anacapa taxonomy tables with Unknown
#' (that is what they effectively are to most users)
#' @param taxon_table taxonomy table from Anacapa
#' @return ana_taxon_table with scrubbed 'sum.taxonomy' column
#' @examples
#' good_taxon_table <- data.frame(sum.taxonomy = c("a;b;c;d;f;u", "p;q;r;s;t;u"),
#' site_1 = c(0,1), site_2 = c(10, 20))
#' scrub_taxon_paths(good_taxon_table)
#' @export
scrub_taxon_paths <- function(taxon_table) {
  to_return <- taxon_table
  new_sum_tax <- reshape2::colsplit(taxon_table$sum.taxonomy, ";",
                          names = c("Phylum", "Class", "Order", "Family", "Genus", "Species"))

  new_sum_tax <- new_sum_tax %>%
    dplyr::mutate(Phylum = ifelse(is.na(Phylum) | Phylum == "", "unknown", Phylum)) %>%
    dplyr::mutate(Class = ifelse(is.na(Class) | Class == "", "unknown", Class)) %>%
    dplyr::mutate(Order = ifelse(is.na(Order) | Order == "", "unknown", Order)) %>%
    dplyr::mutate(Family = ifelse(is.na(Family) | Family == "", "unknown", Family)) %>%
    dplyr::mutate(Genus = ifelse(is.na(Genus) | Genus == "", "unknown", Genus)) %>%
    dplyr::mutate(Species = ifelse(is.na(Species)| Species == "", "unknown", Species))

  new_sum_tax2 <- paste(new_sum_tax$Phylum,
                        new_sum_tax$Class,
                        new_sum_tax$Order,
                        new_sum_tax$Family,
                        new_sum_tax$Genus,
                        new_sum_tax$Species, sep = ";")
  to_return$sum.taxonomy <- new_sum_tax2
  return(to_return)
}

#' Verify that the input taxon_table file and the input mapping file meets specificationss
#' The function takes one taxon table as its input, and verfies that it meets
#' the expected standards.
#' The standards incude:
#' 1. Column names exist.
#' 2. One of the columns is named "sum.taxonomy"
#' 3. The "xxx_seq_number" column, if it ever existed, is removed
#' 4. All columns apart from sum.taxonomy should be numeric
#' 5. All columns apart from sum.taxonomy should have corresponding row in metadata file
#' @param taxon_table taxonomy table from Anacapa
#' @param metadata_file Qiime-style mapping
#' @examples
#' good_taxon_table <- data.frame(sum.taxonomy = c("a;b;c;d;f;u", "p;q;r;s;t;u"),
#' site_1 = c(0,1), site_2 = c(10, 20))
#' good_maps <- data.frame(site = c("site_1", "site_2"),
#' season = c("wet", "dry"), host = c("oak", "sage"))
#' validate_input_files(good_taxon_table, good_maps)
#' @export
validate_input_files <- function(taxon_table, metadata_file) {

  # 1. Column names exist.
  if (is.null(colnames(taxon_table))) {
    stop("The input taxon table should have column names. The taxonomy column should be named 'sum.taxonomy'; the rest of the columns should be named according to their sample names.")
  }

  # 2. One of the columns is named "sum.taxonomy"
  if (!("sum.taxonomy" %in% colnames(taxon_table))) {
    stop("Please make sure that the taxonomy column in the input taxon table is named 'sum.taxonomy'!")
  }

  # 3. The "xxx_seq_number" column, if it ever existed, is removed
  if (any(stringr::str_detect(colnames(taxon_table), "seq_number"))) {
    stop("Please makes sure that you have removed the 'xxx_seq_number' column from the taxon table (note: this can be done with the function `scrub_seqNum_column`)")
  }

  # 4. All columns apart from sum.taxonomy should be numeric
  if (!(all(sapply(taxon_table %>% dplyr::select(-sum.taxonomy), is.numeric)))) {
    stop("Please make sure that all columns apart from sum.taxonomy only contain numeric data!")
  }

  # 5. All columns apart from sum.taxonomy should have corresponding row in metadata file
  if (!(all(colnames(taxon_table %>% dplyr::select(-sum.taxonomy)) %in% metadata_file[, 1]))) {
    stop("Please make sure that each sample in your taxon table has a corresponding row in the mapping file!")
  }


}

```

#Create Phyloseq Objects
```{r}
#Create phyloseq objects

#All Species
physeq_obj.ind <- convert_anacapa_to_phyloseq(t_index, metadata)

#Species Monitored by NPS, PISCO, and Reef Check
physeq_obj.monitored <- convert_anacapa_to_phyloseq(monitored, metadata)

#Species Only Monitored by ReefCheck
physeq_obj.reefcheck <- convert_anacapa_to_phyloseq(reefcheck, metadata)

```


```{r}
# Get counts of richness within different taxonomic levels
get_taxa_unique(physeq_obj.ind, "Class")
get_taxa_unique(physeq_obj.ind, "Order")
get_taxa_unique(physeq_obj.ind, "Family")
get_taxa_unique(physeq_obj.ind, "Genus")
get_taxa_unique(physeq_obj.ind, "Species")
get_taxa_unique(physeq_obj.monitored, "Species")
get_taxa_unique(physeq_obj.reefcheck, "Species")

veganComm_CA <- vegan_otu(physeq_obj.ind)
sp1_CA <- specaccum(veganComm_CA)
sp2_CA <- specaccum(veganComm_CA, "random")
sp2_CA
summary(sp2_CA)
```

```{r}
# Load color palettes
col3.gr <- wes_palette(name = "Darjeeling1", n=5, type="discrete")
col2.gr <- wes_palette(name = "Darjeeling2", n=5, type="discrete")
col1.gr <- wes_palette(name = "GrandBudapest1", n=4, type="discrete")
col4.gr <- wes_palette(name = "Royal1", n=4, type="discrete")
col5.gr <- wes_palette(name = "GrandBudapest2", n=4, type="discrete")
col6.gr <- wes_palette(name = "Chevalier1", n=4, type="discrete")
col7.gr <- wes_palette(name = "Moonrise3", n=4, type="discrete")
col8.gr <- wes_palette(name = "Cavalcanti1", n=5, type="discrete")
col9.gr <- wes_palette(name = "Chevalier1", n=4, type="discrete")
col.great <- c(col1.gr[3],col3.gr[1],col1.gr[2],col3.gr[4],col6.gr[2],col6.gr[1],col2.gr[4],col2.gr[2],col5.gr[2],col4.gr[1],col4.gr[3],col7.gr[2],col7.gr[3],col4.gr[3])
col.great2 <- c(col3.gr[5],col2.gr[2],col7.gr[1],col2.gr[4],col8.gr[4],col3.gr[2],col4.gr[1], col6.gr[3], col9.gr[3])
col10= palette(brewer.pal(n=10,name="PuBuGn"))
colfunc<-colorRampPalette(c(col2.gr[2],col3.gr[5],"springgreen",col9.gr[1]))
plot(rep(1,9),col=(colfunc(9)),pch=19,cex=2)
length(col10)
```

#Figure 2. Heat Map of Monitored Species Detected Over Time. Also Figure S1: Heatmap of all taxa
```{r}
taxa_names(physeq_obj.ind) %>% as_tibble() %>% 
  separate(., col=value, into=c("Phylum","Class", "Order", "Family", "Genus", "Species"), sep="\\;",remove = FALSE) %>%
  mutate(., Species = ifelse(Species=="", Genus, Species)) %>% 
  mutate(., Species = ifelse(Species=="", Family, Species)) %>% 
  mutate(., Species = ifelse(Species=="", Order, Species))-> final_namers

# Heat map for monitored species (Fig 2)
#order taxa by abundance over samples
##extract otu_table and make it as a dataframe
abundance_order<-as(otu_table(physeq_obj.monitored), "matrix")
OTUdf = as.data.frame(abundance_order)
#create column to be able to rank abundance over samples and add species so I can extract the correct order
OTUdf$rank<-apply(OTUdf,1,function(x) sum(x > 0))
rank_species<-row.names(OTUdf)
OTUdf$species<-rank_species
# order
order<-OTUdf[order(OTUdf$rank),]
View(order)
# make the correct order a vector
rank<-order$species
rank_order<-order$rank
abundance_plot<-cbind(rank, rank_order)
abundance_plot<-as.data.frame(abundance_plot)
abundance_plot<- abundance_plot %>% separate(rank, c("phylum", "class", "order", "family", "genus", "species"), ";")
abundance_plot$percentage<-(rank_order/36)*100

# Also make this supplementary figrure to visualize detection in samples
jpeg("detection_plot.tiff", width=5000, height=3300, res=600)
percentage_plot <- ggplot(abundance_plot, aes(x = reorder(species, percentage), y = percentage))+
  geom_bar(stat="identity",  width=0.8,  fill="steelblue") + ylim(NA,100) + coord_flip() + theme_classic() + labs(y= "Detection Percentage", x = "Monitored Species") 
percentage_plot
dev.off()

# make zeros NAs for the na.value to be white
physeq_for_heatmap<-physeq_obj.monitored
otu_table(physeq_for_heatmap)[otu_table(physeq_for_heatmap) == 0.0000] <-NA

#make figure
jpeg("ordered_figure2.tiff", width=5300, height=3300, res=600)
monitored_heatmap<- plot_heatmap(physeq_for_heatmap, sample.label="Time_Point", taxa.label="Species", taxa.order=rank, sample.order="Time_Point",low="#66CCFF", high="#000033", na.value="white")+ scale_fill_gradient(name = "eDNA index", low="#66CCFF", high="#000033", na.value="white", breaks=c(0.25, 0.5, 0.75, 1.0))+theme (axis.text.x = element_text(size=8), axis.text.y = element_text(size=8), axis.title.x = element_text(size=20),axis.title.y = element_text(size=20), legend.title = element_text(size=20), legend.text = element_text(size=16), legend.text.align = 0 )
monitored_heatmap$scales$scales[[1]]$name <- "Time Point Sampled"
monitored_heatmap
dev.off()


#Heat Map of all taxa (Figure S2 )
# First I need to order taxa by abundance over samples
##extract otu_table and make it as a dataframe
abundance_order_all<-as(otu_table(physeq_obj.ind), "matrix")
OTUdf_all = as.data.frame(abundance_order_all)
## create column to be able to rank abundance over samples and add species so I can extract the correct order
OTUdf_all$rank_all<-apply(OTUdf_all,1,function(x) sum(x > 0))
rank_species_all<-row.names(OTUdf_all)
OTUdf_all$species<-rank_species_all
## order
order_all<-OTUdf_all[order(OTUdf_all$rank_all),]
View(order_all)
## make the correct order a vector
rank_all<-order_all$species
rank_order_all<-order_all$rank_all
abundance_plot_all<-cbind(rank_all, rank_order_all)
abundance_plot_all<-as.data.frame(abundance_plot_all)
abundance_plot_all<- abundance_plot_all %>% separate(rank_all, c("phylum", "class", "order", "family", "genus", "species"), ";")
abundance_plot_all$percentage_all<-(rank_order_all/36)*100

# Maybe use this figure?
percentage_plot_all <- ggplot(abundance_plot_all, aes(x = reorder(species, percentage_all), y = percentage_all))+
  geom_bar(stat="identity",  width=0.8,  fill="steelblue") + ylim(NA,100) + coord_flip() + theme_classic() + labs(y= "Detection Percentage", x = "Taxa") 
percentage_plot_all


# make zeros NAs for the na.value to be white
physeq_for_heatmap_all<-physeq_obj.ind
otu_table(physeq_for_heatmap_all)[otu_table(physeq_for_heatmap_all) == 0.0000] <-NA

#make the figure
jpeg("FigureS1.jpg", width=5300, height=3300, res=600)
heat_all <- plot_heatmap(physeq_for_heatmap_all, sample.label="Time_Point", taxa.label="Species", taxa.order=rank_all, sample.order="Time_Point",low="#66CCFF", high="#000033", na.value="white") + scale_fill_gradient(name = "eDNA index", low="#66CCFF", high="#000033", na.value="white", breaks=c(0.25,0.5,0.75,1.0)) + theme (axis.text.x = element_text(size=8), axis.title.x = element_text(size=20),axis.title.y = element_text(size=20), legend.title = element_text(size=20), legend.text = element_text(size=16), legend.text.align = 0 ) 
heat_all$scales$scales[[1]]$name <- "Time Point Sampled"
heat_all
dev.off()



#only Reef Check monitored species 
jpeg("heatmap_monitored.tiff", width=4500, height=3300, res=600)
p<-plot_heatmap(physeq_obj.reefcheck, sample.label="Time_Point", taxa.label="Species", sample.order="Time_Point",low="#66CCFF", high="#000033", na.value="white", trans = log_trans(base = 1.5))+ scale_fill_gradient(name = "eDNA index", low="#66CCFF", high="#000033", na.value="white", breaks=waiver(),trans = log_trans(base = 1.5)) 
p$scales$scales[[1]]$name <- "Time Point Sampled"
p
dev.off()
```
#Figure 3

```{r}
#Generate Vegan formatted data table
sampledf_2 <- data.frame(sample_data(physeq_obj.ind))
carnivore_rel_abun_2<- vegan_otu(physeq_obj.ind)

#Bray curtis dissimilarity matrix
d_carn_2 <- vegdist(carnivore_rel_abun_2, method="bray") 

#Permanova
broom::tidy(adonis(carnivore_rel_abun_2 ~ sampledf_2$Direction+sampledf_2$Replicate+sampledf_2$Time_Point, method = "bray")$aov.tab)

```

```{r}
#Permanova d and dissimMethod for Jaccard and d2 and dissimMethod2 for Bray-Curtis
# Beta diversity stats #how far apart are the clusters (distance)
print(pairwise_adonis(carnivore_rel_abun_2,getElement(sampledf_2, "Direction")))
print(pairwise_adonis(carnivore_rel_abun_2,getElement(sampledf_2, "Replicate")))
print(pairwise_adonis(carnivore_rel_abun_2,getElement(sampledf_2, "Time_Point")))

#Homogeneity of dispersions test
betadisper(d_carn_2, getElement(sampledf_2, "Time_Point"))
broom::tidy(TukeyHSD(betadisper(d_carn_2, getElement(sampledf_2, "Time_Point"))))

betadisper(d_carn_2, getElement(sampledf_2, "Direction"))
broom::tidy(TukeyHSD(betadisper(d_carn_2, getElement(sampledf_2, "Time_Point"))))

betadisper(d_carn_2, getElement(sampledf_2, "Replicate"))
broom::tidy(TukeyHSD(betadisper(d_carn_2, getElement(sampledf_2, "Time_Point"))))
```

```{r}
#Generate Vegan formatted data table
sampledf2_monitored <- data.frame(sample_data(physeq_obj.monitored))
carnivore_rel_abun_monitored <- vegan_otu(physeq_obj.monitored)

#Bray curtis dissimilarity matrix
d_carn_mon <- vegdist(carnivore_rel_abun_monitored, method="bray") 

#Permanova
broom::tidy(adonis(carnivore_rel_abun_monitored ~ sampledf2_monitored$Direction+sampledf2_monitored$Replicate+sampledf2_monitored$Time_Point, method = "bray")$aov.tab)

```


```{r, results="hide",warning=FALSE}
#Apportioning variance: OTU counts

OTU.count.adonis<-adonis(carnivore_rel_abun_2 ~ sampledf_2$Direction+sampledf_2$Replicate+sampledf_2$Time_Point, method = "bray")$aov.tab
OTU.count.adonis

#Set up for Q Plot
q<-as.data.frame(OTU.count.adonis)
row.names(q)<-c("Direction of Tides","Location","Time Point", "Residuals", "Total") # must match names correctly
q<-data.frame(row.names(q), q)
names(q)[1]<-"Level"
q<-q[-which(row.names(q)=="Total"),]
q$Level<-paste(q$Level, "\n",
               round(q$R2,2), "\n"," (p=", q$Pr..F., ")", 
               sep="")

q$Level[4] <- paste("Residuals\n0.54")

col.great2 <- c(col3.gr[2],col2.gr[2],col_Zissou1[1],col3.gr[5])
categories=c("Residual","Time Point","Direction of Tide","Location")
vals_max <- data.frame(pct=factors_max, cat=categories, col=col.great2)
treemap(vals_max, index="cat", vSize="pct", vColor="col", type="color", fontsize.labels = 25)


#Plot Figure 
jpeg("Figure3.jpg",  width=6500, height=4000, res=600)
q0.plot<-ggplot(q, aes(area = R2, label=Level))+
  geom_treemap(fill=col.great2, color=1, alpha=.8)+
  geom_treemap_text(colour = "white", place = "centre", grow = FALSE, size=24, fontface="bold")+
  ggtitle("Apportioned Variance")+
  theme(plot.background = element_blank())+
  theme(plot.title = element_text(hjust = 0.5)) +
   theme(plot.title = element_text(size = 30, face = "bold"),
                       axis.title = element_text(size = 30, face = "bold"),axis.text=element_text(size=24))
q0.plot
dev.off()
```


#Look into detection of species richness with only spatial replicates vs only temporal replicates
```{r}
#Convert to iNEXT format from Observed Species (incidence frequencies)
physeq_obj.noncontam_no_lab.binary <- transform_sample_counts(physeq_obj.ind, function(abund) 1*(abund>0))
physeq_obj.monitored.binary <- transform_sample_counts(physeq_obj.monitored, function(abund) 1*(abund>0))

#subset all taxa
site_A<-subset_samples(physeq_obj.noncontam_no_lab.binary, Replicate == "A")
site_B<-subset_samples(physeq_obj.noncontam_no_lab.binary, Replicate == "B")
site_C<-subset_samples(physeq_obj.noncontam_no_lab.binary, Replicate == "C")
Time0<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T00")
Time1<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T01")
Time2<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T02")
Time3<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T03")
Time4<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T04")
Time5<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T05")
Time6<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T06")
Time7<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T07")
Time8<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T08")
Time9<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T09")
Time10<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T10")
Time11<-subset_samples(physeq_obj.noncontam_no_lab.binary, Time_Point == "T11")

#subset monitored species
site_A_mon<-subset_samples(physeq_obj.monitored.binary, Replicate == "A")
site_B_mon<-subset_samples(physeq_obj.monitored.binary, Replicate == "B")
site_C_mon<-subset_samples(physeq_obj.monitored.binary, Replicate == "C")
Time0_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T00")
Time1_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T01")
Time2_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T02")
Time3_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T03")
Time4_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T04")
Time5_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T05")
Time6_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T06")
Time7_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T07")
Time8_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T08")
Time9_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T09")
Time10_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T10")
Time11_mon<-subset_samples(physeq_obj.monitored.binary, Time_Point == "T11")

# convert to vegan format
veganComm_A <- vegan_otu(site_A)
veganComm_B <- vegan_otu(site_B)
veganComm_C <- vegan_otu(site_C)
veganComm_A_mon <- vegan_otu(site_A_mon)
veganComm_B_mon <- vegan_otu(site_B_mon)
veganComm_C_mon <- vegan_otu(site_C_mon)
vegan_Time0<-vegan_otu(Time0)
vegan_Time1<-vegan_otu(Time1)
vegan_Time2<-vegan_otu(Time2)
vegan_Time3<-vegan_otu(Time3)
vegan_Time4<-vegan_otu(Time4)
vegan_Time5<-vegan_otu(Time5)
vegan_Time6<-vegan_otu(Time6)
vegan_Time7<-vegan_otu(Time7)
vegan_Time8<-vegan_otu(Time8)
vegan_Time9<-vegan_otu(Time9)
vegan_Time10<-vegan_otu(Time10)
vegan_Time11<-vegan_otu(Time11)
vegan_Time0_mon<-vegan_otu(Time0_mon)
vegan_Time1_mon<-vegan_otu(Time1_mon)
vegan_Time2_mon<-vegan_otu(Time2_mon)
vegan_Time3_mon<-vegan_otu(Time3_mon)
vegan_Time4_mon<-vegan_otu(Time4_mon)
vegan_Time5_mon<-vegan_otu(Time5_mon)
vegan_Time6_mon<-vegan_otu(Time6_mon)
vegan_Time7_mon<-vegan_otu(Time7_mon)
vegan_Time8_mon<-vegan_otu(Time8_mon)
vegan_Time9_mon<-vegan_otu(Time9_mon)
vegan_Time10_mon<-vegan_otu(Time10_mon)
vegan_Time11_mon<-vegan_otu(Time11_mon)

# calculate averages for temporal replicates for monitored species only
monitored_temporal<-list ("A"=t(veganComm_A_mon),
                "B"=t(veganComm_B_mon),
                "C"=t(veganComm_C_mon) )
mon_temporal <-lapply(monitored_temporal,  as.incfreq)
t2 <- seq(1,12, by = 1)
out.mon.temporal <- iNEXT(mon_temporal, q=0, datatype="incidence_freq", size=t2)
out.mon.temporal$iNextEst
# 3 samples at A gives average of 25.1 species (23.0-27.1)
# 3 samples at B gives average of 25.6 species (23.7-27.5)
# 3 samples at C gives average of 25.2 species (23.0-27.3)
temporal.mean.mon<-c(25.1,25.6,25.2)
mean(temporal.mean.mon) # 25.3
temporal.min.mon<-c(23, 23.7, 23)
mean(temporal.min.mon) # 23.2
temporal.max.mon<-c(27.1, 27.5, 27.3)
mean(temporal.max.mon) # 27.3

# calculate averages for spatial replicates for monitored species only
monitored_spatial<- list ("T00"=t(vegan_Time0_mon),
                "T01"=t(vegan_Time1_mon),
                "T02"=t(vegan_Time2_mon),
                "T03"=t(vegan_Time3_mon),
                "T04"=t(vegan_Time4_mon),
                "T05"=t(vegan_Time5_mon),
                "T06"=t(vegan_Time6_mon),
                "T07"=t(vegan_Time7_mon),
                "T08"=t(vegan_Time8_mon),
                "T09"=t(vegan_Time9_mon),
                "T10"=t(vegan_Time10_mon),
                "T11"=t(vegan_Time11_mon))
mon_spatial<-lapply(monitored_spatial,  as.incfreq)
t3 <- seq(1,3, by = 1)
out.mon.spatial <- iNEXT(mon_spatial, q=0, datatype="incidence_freq", size=t3)
out.mon.spatial$iNextEst
# 3 samples with T00 gives 26 species (22.6-29.4)
# 3 samples with T01 gives 20 species (18.0-22.0)
# 3 samples with T02 gives 24 species (20.3-27.7)
# 3 samples with T03 gives 26 species (21.4-30.6)
# 3 samples with T04 gives 28 species (24.5-31.5)
# 3 samples with T05 gives 22 species (18.7-25.3)
# 3 samples with T06 gives 19 species (15.7-22.2)
# 3 samples with T07 gives 23 species (19.8-26.2)
# 3 samples with T08 gives 23 species (19.5-26.5)
# 3 samples with T09 gives 23 species (20.5-25.5)
# 3 samples with T10 gives 27 species (22.6-31.4)
# 3 samples with T11 gives 29 species (24.5-33.5)

#mean of spatial replicates
spatial.mean.mon<-c(26, 20, 24, 26, 28, 22, 19, 23, 23, 23, 27, 29)
mean(spatial.mean.mon) # 24.2
lower.spatial.mon<-c(22.6,18,20.3,21.4,24.5, 18.7, 15.7, 19.8, 19.5, 20.5, 22.6, 24.5)
mean(lower.spatial.mon) # 20.7
upper.spatial.mon<-c(29.4, 22.0, 27.7, 30.6, 31.5, 25.3, 22.2, 26.2, 26.5, 25.5, 31.4, 33.5)
mean(upper.spatial.mon) #27.7

# calculate averages for temporal replicates for all taxa
temporal<-list ("A"=t(veganComm_A),
                "B"=t(veganComm_B),
                "C"=t(veganComm_C) )
temporal <-lapply(temporal,  as.incfreq)
out.temporal <- iNEXT(temporal, q=0, datatype="incidence_freq", size=t2)
out.temporal$iNextEst
# 3 samples at A gives average of 39.5 species (36.5-42.4)
# 3 samples at B gives average of 48.5 species (44.8-52.2)
# 3 samples at C gives average of 43.3 species (39.6-46.9)
temporal.mean<-c(39.5,48.5,43.3)
mean(temporal.mean) # 43.8
temporal.min<-c(36.5, 44.8, 39.6)
mean(temporal.min) # 40.3
temporal.max<-c(42.4, 52.2, 46.9)
mean(temporal.max) # 47.2

# calculate averages for spatial replicates for all taxa
spatial<- list ("T00"=t(vegan_Time0),
                "T01"=t(vegan_Time1),
                "T02"=t(vegan_Time2),
                "T03"=t(vegan_Time3),
                "T04"=t(vegan_Time4),
                "T05"=t(vegan_Time5),
                "T06"=t(vegan_Time6),
                "T07"=t(vegan_Time7),
                "T08"=t(vegan_Time8),
                "T09"=t(vegan_Time9),
                "T10"=t(vegan_Time10),
                "T11"=t(vegan_Time11))
spatial<-lapply(spatial,  as.incfreq)
out.spatial <- iNEXT(spatial, q=0, datatype="incidence_freq", size=t3)
out.spatial$iNextEst
# 3 samples with T00 gives 44 species (38.8-49.1)
# 3 samples with T01 gives 39 species (33.9-44.1)
# 3 samples with T02 gives 40 species (31.9-48.1)
# 3 samples with T03 gives 41 species (33.9-48.1)
# 3 samples with T04 gives 45 species (40.1-49.9)
# 3 samples with T05 gives 65 species (55.2-74.8)
# 3 samples with T06 gives 33 species (27.6-38.4)
# 3 samples with T07 gives 32 species (28.2-35.8)
# 3 samples with T08 gives 31 species (25.9-36.1)
# 3 samples with T09 gives 35 species (29.9-40.1)
# 3 samples with T10 gives 47 species (42.1-51.9)
# 3 samples with T11 gives 41 species (33.6-48.4)

#mean of spatial replicates
spatial.mean<-c(44, 39, 40, 41, 45, 65, 33, 32, 31, 35, 47, 41)
mean(spatial.mean) # 41.1
lower.spatial<-c(38.8,33.9,31.9,33.9,40.1, 55.2, 27.6, 28.2, 25.9, 29.9, 42.1, 33.6)
mean(lower.spatial) # 35.1
upper.spatial<-c(49.1, 44.1, 48.1, 48.1, 49.9, 74.8, 38.4, 35.8, 36.1, 40.1, 51.9, 48.4)
mean(upper.spatial) #47.1


```


#Figure 4. eDNA Metabarcoding Species accumulation curves.
```{r}
#Convert to iNEXT format from Observed Species (incidence frequencies)
physeq_obj.reefcheck.binary <- transform_sample_counts(physeq_obj.reefcheck, function(abund) 1*(abund>0))

#Convert to Vegan Dataframe
veganComm2.bin <- vegan_otu(physeq_obj.noncontam_no_lab.binary)
veganComm2.monitored.bin <- vegan_otu(physeq_obj.monitored.binary)
veganComm2.reefcheck.bin <- vegan_otu(physeq_obj.reefcheck.binary)


#Create List for iNEXT Species Incidence Frequencies
mpa_inc <-list ("All Detected"=t(veganComm2.bin),
                "All Monitored"=t(veganComm2.monitored.bin),
                "Reef Check"=t(veganComm2.reefcheck.bin) )

#Convert to iNext formatted object
species_incidence <-lapply(mpa_inc, as.incfreq)
View(species_incidence)
# Set Sampling Steps
t <- seq(1, 36, by=1)

#Run iNEXT format
out.inc <- iNEXT(species_incidence, q=0, datatype="incidence_freq", size=t)

#View Species Capture Estimates
out.inc$iNextEst
out.inc$DataInfo

# Sample‐size‐based R/E curves
jpeg("Figure4.jpg",  width=6500, height=4000, res=600)
ggiNEXT(out.inc, type=1) + 
  facet_wrap(~site, scales="free") + aes(fill=site, color = site) +
  theme_bw(base_size = 18) + scale_fill_manual(values=col_species, name = "Species List") + scale_color_manual(values=col_species, "Species List") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank()) + ggtitle("") + xlim(1,40) +scale_x_continuous(breaks=c(0,3,6,12,18,24,30,36)) + xlab("Number of Samples") +ylab("Number of Species") +
   theme(plot.title = element_text(size = 20, face = "bold"),
                       axis.title = element_text(size = 20, face = "bold"),axis.text=element_text(size=14), legend.title = element_text(size = 20, face = "bold") , legend.text = element_text(size=14) )+
  scale_y_continuous(breaks = scales::pretty_breaks(7), limits = c(0, NA)) + theme(strip.text.x = element_text(size = 16, colour = "black", face = "bold.italic")) +  guides(linetype=FALSE,
           colour=FALSE, 
           fill=FALSE, 
           shape=FALSE)
dev.off()

```

#Supplemental Figures
```{r}
#NMDS (Figure S2A and Figure S2B )
ord_all_2020 <- ordinate(physeq_obj.ind, method = "NMDS", distance = d_carn_2)
sample_data(physeq_obj.ind)
col.four <- c(col_darjeeling1[3],col_darjeeling1[1],col_darjeeling1[2],col_Darjeeling2[2])
jpeg("NMDS_DirectionTide_allCA.jpg", width=4000, height=3300, res=600)
nmdsplot_max <- plot_ordination(physeq_obj.ind, ord_all_2020, "Direction", color = "Direction", shape = "Direction") +
  geom_point(size=5) +
  theme_bw() +  stat_ellipse(type = "t", geom = "polygon", alpha = 0.2, aes(group=Direction, fill=Direction)) +
  scale_colour_manual(values=col_nmds, name="Direction of Tide",
                      labels=c("Outgoing Tide","Peak Tide","Incoming Tide")) +
  scale_fill_manual(values=col_nmds, name="Direction of Tide",
                      labels=c("Outgoing Tide","Peak Tide","Incoming Tide"))+
  scale_shape_manual(values=c(15,16,17,18,19), name="Direction of Tide",
                      labels=c("Outgoing Tide","Peak Tide","Incoming Tide")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, face="bold"))
nmdsplot_max  
dev.off()

#only CA species data time NMDS Bray
jpeg("NMDS_TimePoint_allCA.tiff", width=4000, height=3300, res=600)
nmdsplot_max_time <- plot_ordination(physeq_obj.ind, ord_all_2020, "Time_Point", color = "Time_Point") +
  geom_point(size=5) +
  theme_bw() +  geom_polygon(type = "t", geom = "polygon", alpha = 0.2, aes(group=Time_Point, fill=Time_Point)) +
  scale_color_manual(values=col.great)+
  scale_fill_manual(values=col.great)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, face="bold")) + labs(fill = "Time Point", color="Time Point")
nmdsplot_max_time
dev.off()

#monitored species data tide NMDS Bray
nmdsplot_monitored <- plot_ordination(physeq_obj.step2_species_monitored, ord_monitored, "High.Low", color = "High.Low", shape = "High.Low") +
  geom_point(size=5) +
  theme_bw() +  stat_ellipse(type = "t", geom = "polygon", alpha = 0.2) +
  theme_ranacapa() +
  scale_color_manual(values=col.great)
nmdsplot_monitored
```

```{r}
#plot the averages for spatial vs temporal replicates
timevsspace<-data.frame(Number_of_Samples, SpatialReplicates, TemporalReplicates)
timevsspace2 <- melt(timevsspace, id.vars='Number_of_Samples')
timevsspace2$se<-SE
plot_timevsspace<-ggplot(data=timevsspace2, aes(x=Number_of_Samples, y=value, fill=variable)) +
  geom_bar(stat="identity", position = 'dodge')+
  geom_errorbar(aes(ymin=value-SE, ymax=value+SE), width=.2, position=position_dodge(.9))
plot_timevsspace

```

